A Graph Theoretic Approach to Problem Formulation for Multidisciplinary Design Analysis and Optimization

### 2013-09-05 reviewer_response.tex

\rev{Page 16 line 38 
column 2: Typo : “… formulations involving MULTIPLES analysis ….”}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Reviewer 2}

\rev{Page 1, Line 46: 2nd column...check spacing after ``ModelCenter''.}
Fixed

\rev{Page 2, Line 38: 2nd column...as your motivation was really done in the 
Introduction section, consider changing the heading on section 2 to 
``Background'' or something similar.  You don't really provide motivation in 
this section anyhow.}
Section title changed

\rev{Page 3, Line 19, 1st column: change ``weather'' to ``whether''}
fixed

\rev{Page 3, Line 21, 2nd column: change ``descrition'' to ``description''}
fixed

\rev{Page 4, Line 37, 1st column: what does ``REMS'' stand for?}
"Reconfigurability in MDO Problem Synthesis". Full name added to the paper.

\rev{Page 4, Line 45, 1st column: change ``REMS address'' to ``REMS addresses''}
fixed

@@

\rev{Page 4, Lines 35 to 55, 1st column:  This paragraph is a bit awkward in 
structure.  In the second to last sentence you state, "REMS provides syntax 
for variables and functions but does not facilitate inclusion of solvers or 
optimizers in the graph."  The reader then expects your method to address 
these shortcomings in the next sentence but instead you state, "...the goal 
of allowing graphs to describe analysis couplings and more closely interact 
with existing design processes and MDAO frameworks," which you never really 
called out as a shortcoming of REMS in the first place.  I think you could 
improve the flow/wording here.}

This paragraph was poorly worded. We have re-worded it to attempt 
to make the issue with REMS more clear and to address more bluntly 
what it fails to allow.

@@

\rev{Line 10, 1st column, Page 10:  This equation (missing a number) has a LOT of 
unions, variables, etc...in it.  It would help tremendously to provide a 
verbal description of at least one of these equations to help with 
translation into the real world.  This would be similar to what you did 
following the steps (1) to (4) in the next subsection.}

equation numbers were added, see new equations 8 and 9. Also, verbal descriptions
were added to accompany the set notations.

@@

\rev{Lines 5 - 12, 2nd column, Page 12:  Just a general question spurred by the 
text here.  What criteria would be applied to determine which FPG is better 
than another?}

We chose not to address the issue of seleting one FPG over another. The answer 
to this quetion depends heavily on specific situation. If you're short on time, 
a smaller FPG with less cycles and less variables may be preferable. If you're trying
to resolve some small issue with nacelle-wing interference then you may choose to go with a
more high fidelity and computationally expensive FPG.

Our main point was that the MCG provides a good foundation for deriving the FPG, and that 
it can be revisited later in the design process as things change. You are not stuck modifying the 
FPG you initally select.

\rev{Line 57, 1st column, Page 13: change ``as a input'' to ``as an input''}
fixed

@@

\rev{Now for the request of more significant importance.  This is coming from 
someone with significant experience in MDO and who uses a lot of the 
analysis tools you specified in your case study on a daily basis.  I really 
liked this article but felt it was lacking some punch in the results 
section.  This was namely in the form of quantitative results.  You vaguely 
give some in the form of the different resulting connectivity graphs and 
this was good but to really drive home the value of your research to a 
potential user I think you could do a bit better with very minimal work. 
The potential user of this methodology will be concerned with one of two 
things: 1) minimizing computational time or 2) increasing the accuracy of 
the results.  I think you've given examples of these two scenarios already 
in figures 15 and 16, so why not provide a table of the CPU times (1) and/or 
deviation from higher order analyses (2) for each of these graphs.  Doesn't 
have to necessarily be these two metrics but something to really tie the 
output of your methodology to something tangible for comparison purposes. 
You could even include the metrics for a randomly obtained FPG as a 
baseline.  Then the person reading the article can say, "Wow, given a set of 
analysis tools and a general problem, I can employ this method to obtain an 
FPG that significantly reduces my overall computational time." or "Wow, 
given a set of analysis tools and a general problem, I can employ this 
method to obtain an FPG that significantly improves my overall computational 
accuracy while keeping my computational efficiency as high as possible."}

We agree that this sort of quantitative data would be very valuable. However, we 
contend that this data needs to be larger in scope than just that which could be obtained 
by executing the example problem given here. Instead we think that to do that topic 
justice, you would need to address a wide range of problems with different graph structural 
characteristics. The challenge is that any quantitative measures given are inherantly 
dependent on particular details of a problem. So to address this question fully you 
would need a range of problems to consider. This work, while very valuable, would 
fill an entire paper in it's own right and it's something we will consider in future 
work.

@@

\rev{If you are worried about length, I think the conclusions section may be cut 
in half and still accomplish its intended purpose.}

Althought the length of the paper was not excessive, we agree that the conclusion 
could be more concise. We have shortened the conclusion and re-woreded it to make 
it more clear.

### 2013-09-05 conclusions.tex
addresses second reviewers comments

\section*{Conclusions}
In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory. 
Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, DSM, and XDSM, but it provides new constructs tailored to applications in early phases of MDAO problem formulation before specific problem solution approaches are envisioned. 
In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and the Fundamental Problem Graph (FPG).

The MCG addresses the question, ``Given a set of analysis tools, what are all of the variable interconnections between them that could be established?''  
The MCG provides a structured formalism to identify \textit{holes} and \textit{collisions} in interconnections between input and output variables implied by the set of analysis tools. 
A hole corresponds to a variable that is not calculated in the MCG data flow but is needed by other tools. 
To create a valid problem formulation, all holes must be ``filled'' either by assigning the corresponding variable as a design variable to be set by the user or an optimizer, or by introducing additional tools to compute the variable.

A collision corresponds to a situation in which two or more analysis tools compute separate instances of the same variable. 
To create a valid problem formulation, all collisions must be resolved either by removing redundant analysis tools or by introducing logic or solvers into the data flow to choose between or aggregate multiple instances during execution. 
The identification of collisions in an MCG is especially important for problem formulations involving multiple analysis tools of different fidelities; in this context, a collision represents an opportunity to formulate a multi-fidelity problem that can take best advantage of the efficiency/accuracy tradeoff between different analysis tools. 
Resolving collisions and filling holes in an MCG requires choices by the user to specify the overall MDAO problem that he/she desires to solve, i.e. to produce the required set of system-level outputs given a specified set of design variables. 
One benefit of a graph theoretic approach is the standard algorithms which can be used to inform the user's decisions, such as cycle detection, minimum spanning trees, and shortest path algorithms.

The FPG is a graph that describes a data connection structure corresponding to an MDAO problem formulation free of collisions and holes. 
An FPG is the result of user choices to fill holes and resolve collisions in the initial MCG. 
Typically, many different FPGs could be attained from a given MCG, depending on the user choices. 
The number of possible FPGs that can be attained by selecting different design variables or introducing additional analysis tools to fill holes and/or resolve collisions can be viewed as a measure of the freedom available to the user in implementing the available analysis tools to formulate valid MDAO problems. 
In Sec.~\ref{s:building graphs}, we present an algorithm for identifying holes and conflicts in an MCG that can be applied by the user in a process to arrive at valid FPGs. 
In Sec.~\ref{s:example problem}, we provide an example application of formulating an FPG from an MCG for a commercial aircraft design problem.

Once a desired FPG is identified, the graph syntax described in the paper can be leveraged to formulate a Problem Solution Graph (PSG). The PSG describes particular MDAO solution architectures, e.g. MDF, IDF, BLISS, ATC, that can be used to solve the problem defined by the FPG. 
Many possible PSGs could be formulated to solve a problem corresponding to a specified FPG.  
The application of the syntax to PSGs has not been illustrated in the paper.

For simple problems with few analysis tools and variables, the formulation of a valid problem description is straightforward. 
However, MDAO problems continue to grow in scale and complexity. 
As the numbers of analysis tools and variables have increased, it has become increasingly challenging and time consuming for engineering teams to determine how the multiple analysis tools can be interconnected to produce valid problem formulations, to know when other tools must be introduced, and to determine the number of free variables in the problem that can or should be varied by the designer or an optimizer. 
The graph formalism presented in this paper is intended to offer value in this context of increasing problem complexity by providing a formal approach to identify and resolve missing and redundant data in order to create valid MDAO problem formulations.

>>>>

\section*{Conclusions}
In this paper, we have presented an approach for describing MDAO problems with constructs and syntax from graph theory. 
Our graph description shares similarities to other approaches such as REMS, $\Psi$, FDT, 
DSM, and XDSM, but it provides new constructs tailored to algorithmic analysis and 
manipulation. The proposed syntax has applications in early phases of MDAO problem 
formulation, especially for problems with large numbers of variables and analyses. 
In particular, we introduce the concepts of the Maximally Connected Graph (MCG) and 
the Fundamental Problem Graph (FPG).

The MCG addresses the question, ``Given a set of analysis tools, 
what are all of the variable interconnections between them that could be established?''  
The MCG provides a structured formalism to identify \textit{holes} and 
\textit{collisions} in interconnections between the set of analysis tools. 
In order to achieve a valid problem formulation, all of the holes and collisions must 
be resolved. We present an algorithm to perform this resolution which gurantees that you 
will find an FPG if one exists. In some cases this algorithm still relies on some input from the 
designers to make decisions. One benefit of a graph theoretic approach is the standard 
algorithms which can be used to inform the user's decisions, such as cycle detection, 
minimum spanning trees, and shortest path algorithms.

The FPG is a graph that describes a data connection structure corresponding 
to an MDAO problem formulation free of collisions and holes. 
An FPG is the result of user choices to fill holes and resolve collisions in 
the initial MCG. Typically, many different FPGs could be attained from a given MCG, 
depending on the user choices. The number of possible FPGs that can be attained by 
selecting different design variables or introducing additional analysis tools to 
fill holes and/or resolve collisions can be viewed as a measure of the freedom 
available to the user in implementing the available analysis tools to 
formulate valid MDAO problems. In Sec.~\ref{s:example problem}, 
we provide an example application of formulating multiple FPGs from an MCG 
for a commercial aircraft design problem.

For simple problems with few analysis tools and variables, the formulation of a 
valid problem description is straightforward. However, MDAO problems continue to 
grow in scale and complexity. As the numbers of analysis tools and variables have 
increased, it has become increasingly challenging and time consuming for engineering 
teams to determine how the multiple analysis tools can be interconnected to produce 
valid problem formulations, to know when other tools must be introduced, and to 
determine the number of free variables in the problem that can or should be 
varied by the designer or an optimizer. The graph formalism presented in this 
paper is intended to offer value in this context of increasing problem complexity by 
providing a formal approach to programatically identify and resolve missing and 
redundant data in order to create valid MDAO problem formulations.

### 2013-09-05 intro.tex
addresses second reviewers comments

    The time and expense for setting up 
    analysis models is growing in conjunction with the increasing scope of MDO problems. 
    Multidisciplinary Design Analysis and Optimization (MDAO)
    frameworks such as {ModelCenter}\textregistered, iSigt\textregistered, ModeFrontier\textregistered, and OpenMDAO \cite{Gray2012}and have enabled a new level of analysis tool integration 
    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
    new capability has created a new challenge. Even models with tens of analyses could include hundreds or thousands
    of variables that are interdependent and must be linked in the framework. 
    A common situation is that different analyses provide differing values for the 
    same physical quantity, often in different data formats, and these conflicts need to be resolved. 
    These occurrences are particularly acute in situations in which analysis tools have differing fidelities. For
    example, an abstracted aerodynamic analysis such as an empirical drag buildup model may return only
    integrated drag, whereas a CFD tool may return pressure and shear stress distributions across the entire
    surface grid. If an analysis downstream of the aerodynamics tool needs only integrated drag as an input,
    then the designer has a free choice of which of the two possible aerodynamics analysis tools to select to
    provide the drag estimate (presuming that drag can be computed from surface distributions by a simple
    integration algorithm). On the other hand, if a downstream analysis needs a pressure distribution in
    order to compute pitching moment, for example, then any feasible MDO problem formulation must
    include CFD or similar analysis in the data flow, regardless of whether the empirical drag buildup model
    is also included.

@@

\section{Motivation}
\subsection{Specific vs Fundamental Problem Formulation }
	\label{s:specific vs fundamental}
    The Fundamental Problem Formulation (FPF) is our terminology for a statement of the overall system-level optimization problem that contains only  information about analysis tools, design variables, objectives, and constraints without reference to a solution approach. 
In particular, the FPF does not depend on the choice of solution-specific elements such as the MDAO framework, optimization algorithm, iterative solver, or execution sequence. 
This description represents  the problem from the point of view of an engineer specifying the design problem that he/she desires to solve without specifying how the problem should be solved. 
The lack of solution information in the FPF contrasts with optimization problem statements that are written with reference to a specific solution strategy or with implication of a specific execution sequence. 
For example, consider the Sellar test problem \cite{AIAA:sellar}, with an FPF given as follows:

@@

    In Eq.~\ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
    to $D_2$. Similarly, $y_2$ is an output of $D_2$ as well as an input 
    to $D_1$. This implies that $D_1$ and $D_2$ are coupled together by means 
    of their mutual dependence. However no information is given as to weather
    $D_1$ or $D_2$ should be run before the other. Given Eq.~\ref{eqn:simple_fpf}, 
    it would be equally valid to run $D_1$ first, $D_2$ first, or both simultaneously. 
Any MDAO root finding method or compatibility constraint formulation could be implemented to achieve consistency between $D_1$ and $D_2$.

@@

    Alexandrov and Lewis introduced a graph based syntax called REMS which 
    incorporates objectives and constraints into a graph, effectively combining 
    FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
    matrix form from DSM, but by adding the objectives and constraints, it partially 
    combines a traditional DSM with an FDT. This formulation allows REMS to represent data 
    dependency between multiple analysis tools as well as between analysis tools and
    objective/constraint functions. Additionally, REMS address the need to
    transition between multiple solution strategies while maintaining a single consistent  
    graph representation of the fundamental problem formulation. REMS provides
    syntax for variables and functions but does not facilitate inclusion 
    of solvers or optimizers in the graph.  The graph syntax proposed in this paper could  
    be viewed as an extension to REMS with the goal of allowing the graphs to  describe analysis couplings and more 
    closely interact with existing design processes and MDAO frameworks.

>>>>

    The time and expense for setting up 
    analysis models is growing in conjunction with the increasing scope of MDO problems. 
    Multidisciplinary Design Analysis and Optimization (MDAO)
    frameworks such as iSight\textsuperscript{\textregistered},\\ ModelCenter\textsuperscript{\textregistered}, ModeFrontier\textsuperscript{\textregistered}, and OpenMDAO\cite{Gray2012} and have enabled a new level of analysis tool integration 
    and have paved the way for models with more analyses and increasing numbers of interdisciplinary couplings. This 
    new capability has created a new challenge. Even models with tens of analyses could include hundreds or thousands
    of variables that are interdependent and must be linked in the framework. 
    A common situation is that different analyses provide differing values for the 
    same physical quantity, often in different data formats, and these conflicts need to be resolved. 
    These occurrences are particularly acute in situations in which analysis tools have differing fidelities. For
    example, an abstracted aerodynamic analysis such as an empirical drag buildup model may return only
    integrated drag, whereas a CFD tool may return pressure and shear stress distributions across the entire
    surface grid. If an analysis downstream of the aerodynamics tool needs only integrated drag as an input,
    then the designer has a free choice of which of the two possible aerodynamics analysis tools to select to
    provide the drag estimate (presuming that drag can be computed from surface distributions by a simple
    integration algorithm). On the other hand, if a downstream analysis needs a pressure distribution in
    order to compute pitching moment, for example, then any feasible MDO problem formulation must
    include CFD or similar analysis in the data flow, regardless of whether the empirical drag buildup model
    is also included.

@@

\section{Background}
\subsection{Specific vs Fundamental Problem Formulation }
	\label{s:specific vs fundamental}
    The Fundamental Problem Formulation (FPF) is our terminology for a statement of the overall system-level optimization problem that contains only  information about analysis tools, design variables, objectives, and constraints without reference to a solution approach. 
In particular, the FPF does not depend on the choice of solution-specific elements such as the MDAO framework, optimization algorithm, iterative solver, or execution sequence. 
This description represents  the problem from the point of view of an engineer specifying the design problem that he/she desires to solve without specifying how the problem should be solved. 
The lack of solution information in the FPF contrasts with optimization problem statements that are written with reference to a specific solution strategy or with implication of a specific execution sequence. 
For example, consider the Sellar test problem \cite{AIAA:sellar}, with an FPF given as follows:

@@

    In Eq.~\ref{eqn:simple_fpf}, $y_1$ is an output of $D_1$ as well as an input 
    to $D_2$. Similarly, $y_2$ is an output of $D_2$ as well as an input 
    to $D_1$. This implies that $D_1$ and $D_2$ are coupled together by means 
    of their mutual dependence. However no information is given as to whether
    $D_1$ or $D_2$ should be run before the other. Given Eq.~\ref{eqn:simple_fpf}, 
    it would be equally valid to run $D_1$ first, $D_2$ first, or both simultaneously. 
Any MDAO root finding method or compatibility constraint formulation could be implemented to achieve consistency between $D_1$ and $D_2$.

@@

    Alexandrov and Lewis introduced a graph based syntax called Reconfigurability in 
    MDO Problem Synthesis (REMS) which incorporates objectives and constraints 
    into a graph, effectively combining FDT and DSM \cite{alexandrov2004}. REMS retains the square adjacency 
    matrix form from DSM, but by adding the objectives and constraints, it partially 
    combines a traditional DSM with an FDT. This formulation allows REMS to represent data 
    dependency between multiple analysis tools as well as between analysis tools and
    objective/constraint functions. Additionally, REMS addresses the need to
    transition between multiple solution strategies while maintaining a single consistent  
    graph representation of the fundamental problem formulation. However REMS does not provide 
    a mechanism for inclusion of solvers or optimizers in the graph. This fundamentally limits REMS 
    from describing the specifics of a different solution strategies as applied to a specific problem.

### 2013-09-05 mcg_fpg.tex
addresses second reviewers comments

    We may then construct the maximal connectivity graph $M=(V_M,E_M)$ as
    \begin{IEEEeqnarray*}{rCl}
    V_M & = & V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right), \\
    E_M & = & E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right).
    \end{IEEEeqnarray*}
    The MCG $M$ is uniquely determined by the given set of analysis blocks, inputs, and expression blocks. 
    In the cases where the set of inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.

>>>>

    We may then construct the maximal connectivity graph $M=(V_M,E_M)$. $V_M$ is 
    the union of the sets of variable nodes from the inputs, analysis blocks, and expression blocks. 
    \begin{equation}
    V_M = V_\txt{in} \cup \left( \bigcup_{i \in I_A} V_{A_i} \right) \cup \left( \bigcup_{i \in I_E} V_{E_i} \right)
    \end{equation}
    $E_M$ is the union of all connection and fixed edges from the all the analysis blocks and expression blocks. 
    \begin{equation}
    E_M = E_{M,\txt{con}} \cup \left( \bigcup_{i \in I_A} E_{A_i} \right)  \cup \left( \bigcup_{i \in I_E} E_{E_i} \right)
    \end{equation}
    The MCG $M$ is uniquely determined by the given set of analysis blocks, inputs, and expression blocks. 
    In the cases where the set of inputs is not known a priori, the process of obtaining the FPG will reveal the required inputs, as discussed subsequently.
